Supplimentary materials

### 2020-08-14 manuscript.Rmd
pre figure removeal

author: 
  - name          : "Harry Fisher"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       :  "Cyncoed Road, Cardiff, CF23 6XD"
    email         : "harryfisher21@gmail.com"
  - name: Dr Marianne Gittoes
    affiliation: '1'
  - name: Professor Lynne Evans
    affiliation: '1'
  - name: Miss Leah Bitchell
    affiliation: '1'  
  - name: Dr Richard Mullen  
    affiliation: '2'
  - name: Dr Marco Scutari
    affiliation: '3'

@@

@@

keywords          : "Sports Injury, Stress, Interdisciplinary, Bayesian Network"

@@

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 source("r-files/functions.R")
 source("r-files/methods_code.R")
 source("r-files/results_code.R")
#load("all_data.RData", envir = myEnv)
#myEnv <- new.env()
#attach(myEnv, name="sourced_scripts")

@@

10422 words in text body
3142 words in reference section

@@

Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005]. Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010].However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.

@@

Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury. However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence. Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014]. Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury. Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.

@@

A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (S1 Table). 
The LESCA comprises 69 items that reflect possible life events that participants may have experienced. Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”. Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4  ($extremely\;negative$) and +4 ($extremely\;positive$). 
Negative and positive life event scores were calculated by summing the negative and positive scores, respectively. A score for total life events was also calculated by summing the absolute values for both negative and positive events. 
Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. Petrie also provided evidence of discriminant, convergent and predictive validity. For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire. The composite reliability for the LESCA in this study was .84.

@@

A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`. Further details regarding the revised RST are in S1 Appendix.

@@

```{r fig3, fig.cap = 'Study design. For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width="100%"}

include_graphics("figures_doc/studydesign_new.png")
```

@@

```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
include_graphics("figures_doc/sessionprotocol.png")
```

@@

Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014]. 
A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [Fig \ref{fig:fig5}; @Pearl1988]. 
However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012]. In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.

```{r fig5, fig.cap="A simple discrete Bayesian network contain nodes, possible states of the nodes and the arcs connecting nodes.", fig.pos="H", out.width="100%", fig.align='center'}

include_graphics("figures_doc/examplenetwork.png")

@@

Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point. 
To prepare the data for the BN, missing values in the dataset were first imputed. 
Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings. The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random. 
A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
In the current study, the _caret_ package [@Kuhn2008] was used to impute the missing values. 
A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference. 
Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist). 
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014]. 
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
Table \ref{tab:table2} provides an example of the formatted data and demonstrates that participants 1 and 3 have complete data, and therefore have two rows of data each representing variables from T1 $\rightarrow$ T2 and T2 $\rightarrow$ T3, respectively. 
Participant 2 did not complete the final data collection at T3 and therefore only has one row of data representing the variables collected at T1 and T2.
In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables. 
Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
All subsequent models used the same blacklist.

```{r table2, results='asis'}
exampledata <- tibble(
  Participant = c("1", "1", "2", "3", "3"),
  "X\\_1" = c("T1", "T2", "T1", "T1", "T2"),
  "X\\_2" = c("T2", "T3", "T2", "T2", "T3")
)

kable(
  exampledata,
  caption = "Example of the data arrangement used for the network.",
  escape = FALSE
) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE)
```

@@

Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (highest BIC value = nleleg, BIC values relative to nlelg; tlelg only = `r le[2,2]-le[1,2]`, tlelg and nlelg = `r le[3,2]-le[1,2]`). 
Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data. 
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure. TLE score was not included in the final structure.

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]- rstmodel[1,2]`), GDP (`r rstmodel[4,2]-  rstmodel[1,2]`) and I (`r rstmodel[5,2] - rstmodel[1,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2] - rstmodel[1,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure. 
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

@@

Preliminary network structures also revealed strong dependencies between the same variables at sequential time points. For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%. Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network. Therefore, these arcs were blacklisted from the network.
To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018]. Arcs that were present in at least 30% of the models were included in the averaged model. 
The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction. An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks. In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013]. 
Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in S2 Table and S3 Table respectively.

@@

Change scores between timepoint 1 and timepoint 2 were standardized to allow relative changes between variables to be compared.

@@

Conditional probability queries (CPQ) were again used to perform inference the network structures.
The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
The samples were then examined using a Bayesian linear regression models with the _brms_ package [@Burkner2017a] to determine the relationship between these nodes. 
Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.

@@

cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
  kable(
    .,
    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
    escape = F
  ) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
  footnote(general = "Other included bone, skin and brain injuries.") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE)

@@

\newpage

```{r fig6, fig.align='default', fig.height=8, fig.width=12, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables. The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
renderGraph(gR)
```

@@

```{r table5, results='asis'}
query1 %>%
  select(c(1, 2, 3)) %>%
  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable.",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in S4 Table).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`

@@

The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in S5 Table).
The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".

@@

Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket.
The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).

@@

## Summary fo results

>>>>

author: 
  - name          : "Harry Fisher"
    affiliation   : "1"
    corresponding : yes    
    address       :  "Cyncoed Road, Cardiff, CF23 6XD"
    email         : "harryfisher21@gmail.com"
  - name: Dr Marianne Gittoes
    affiliation: '1'
  - name: Professor Lynne Evans
    affiliation: '1'
  - name: Miss Leah Bitchell
    affiliation: '1'  
  - name: Dr Richard Mullen  
    affiliation: '2'
  - name: Dr Marco Scutari
    affiliation: '3'

@@

authornote: |
    12961 words in text body and refence section

    Number of tables: 11

    Number of figures: 10

    Format: British English

@@

keywords          : "Sports injury, Stress, Interdisciplinary, Bayesian Network, Sports psychology"

@@

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r include=FALSE}
options(scipen=999, digits = 2)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
 # source("r-files/functions.R")
 # source("r-files/methods_code.R")
 # source("r-files/results_code.R")

load("all_data.RData")
#myEnv <- new.env()
#attach(myEnv, name="sourced_scripts")

@@

@@

Within Williams and Andersen’s [@Williams1998] model, major life events, which is a component of an athlete’s history of stressors, most consistently predicts injury occurrence [@Williams2007]; specifically, major life events with a negative, as opposed to positive, valence [@Passer1983a; @Maddison2005].
Personality traits and coping resources have also been found to predict injury with athletes more likely to sustain an injury if they have poor social support and psychological coping skills, and high trait anxiety and elevated competitive state anxiety; compared to those with the opposing profile [@Smith1990; @Lavallee1996; @Ivarsson2010]. 
However, the amount of variance explained by these psychosocial factors has been modest and typically between 5 - 30% [@Galambos2005; @Ivarsson2010], which indicates a likely interaction with other factors.

@@

Although research supporting the BMSAIH has mainly focused on the relationship between hormonal responses to training and injury occurrence [@Perna1995; @Perna1997; @Perna2003], other training-related stress markers including heart rate variability [@Bellenger2016; @Williams2017],  muscle stiffness [@Pruyn2015] and postural stability [@Romero-Franco2014] have been reported to be associated with an increased risk of injury. However, the largely mono-disciplinary approach to examining the role of these markers has provided limited insight into the multifaceted interactions with psychological factors that may contribute to injury occurrence. 
Furthermore, a reliance on designs that capture a single point of measurement precludes the assessment of intra- as well as inter-individual changes and the effect of the time interval between measurement and injury occurrence on subsequent injuries [@Johnson2014]. Such an approach fails to capture changes in both psychological factors and physiological markers that may occur preceding an injury. 
Importantly, a repeated measurement approach would enable an assessing of how variables and their interactions change with respect to time provide greater insight into the effect that repeated exposure to major life events and other stress-related factors has on injury occurrence.

@@

A modified version of the Life Events Survey for Collegiate Athletes (LESCA) was used to measure participants’ history of life event stress [@Petrie1992]. 
The LESCA is the most widely used measure of major life events for athletes in the sports injury literature. 
Modifications were made to the LESCA to ensure the suitability of the items for the study cohort (supplementary table 1). 
The LESCA comprises 69 items that reflect possible life events that participants may have experienced. Example items include, “Major change in the frequency (increased or decreased) of social activities due to participation in sport”, “Major change in the amount (more or less) of academic activity (homework, class time, etc)” and “Major change in level of athletic performance in actual competition (better or worse)”. Participants were asked to rate the perceived impact of each life event they had experienced within 12-months preceding the study onset on an 8-point Likert scale anchored at -4  ($extremely\;negative$) and +4 ($extremely\;positive$). 
Negative and positive life event scores were calculated by summing the negative and positive scores, respectively. A score for total life events was also calculated by summing the absolute values for both negative and positive events. 
Petrie (1992) [@Petrie1992] reported test-retest reliabilities at 1-week and 8-weeks with values ranging from .76 to .84 (p < .001) and .48 to .72 (p < .001) respectively. Petrie also provided evidence of discriminant, convergent and predictive validity. For this study, composite reliability [@Fornell1981] was preferred to Cronbach’s alpha as it does not assume parallelity (i.e., all factor loadings are constrained to be equal, and all error variances are constrained to be equal) and instead takes into consideration the varying factor loadings of the items in the questionnaire. The composite reliability for the LESCA in this study was .84.

@@

A revised version of the Reinforcement Sensitivity Theory Personality Questionnaire (RST-PQ) was used to measure motivation, emotion, personality and their relevance to psychopathology [@Corr2016c]. 
The revised version of the RST-PQ comprises 51 statements that measure three major systems: Fight-Flight-Freeze System (FFFS; e.g., "I am the sort of person who easily freezes-up when scared"), Behavioural Inhibition System (BIS; e.g., "When trying to make a decision, I find myself constantly chewing it over") and four Behavioural Approach System (BAS) factors; Reward Interest (e.g., “I regularly try new activities just to see if I enjoy them”), Goal Drive Persistence (e.g., “I am very persistent in achieving my goals”), Reward Reactivity (e.g., “I get a special thrill when I am praised for something I've done well”) and Impulsivity (e.g., “I find myself doing things on the spur of the moment”).
Participants rated each item on a scale from 1 ($not\; at\; all$) to 4 ($highly$) to reflect how well each statement described their personality in general. 
The responses to items associated with each subscale (FFFS, BIS, RI, GDP, RR and I) were summed to give a total personality score that was subsequently used for further analysis.
The composite reliabilities for each subscale were; BIS = `r rst_rel$final[1]`, FFFS = `r rst_rel$final[2]`, GDP = `r rst_rel$final[3]`, I = `r rst_rel$final[4]`, RI = `r rst_rel$final[5]`, RR = `r rst_rel$final[6]`. Further details regarding the revised RST are in supplementary Appendix 1.

@@

```{r fig3, fig.cap = 'Study design. For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width = "100%"}

include_graphics("figures_doc/Fig3.jpeg")
```

@@

```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
include_graphics("figures_doc/Fig4.jpeg")
```

@@

Two Bayesian Networks (BN) were used to explore the relationships between the psychological measures, physiological markers of stress and sports injury. 
A BN is a graphical representation of a joint probability distribution among a set of random variables, and provides a statistical model describing the dependencies and conditional independences from empirical data in a visually appealing way [@Scutari2014]. 
A BN consists of arcs and nodes that together are formally known as a directed acyclic graph (DAG), where a node is termed a parent of a child if there is an arc directed from the former to the latter [@Pearl1988]. 
However, the direction of the arc does not necessarily imply causation, and the relationship between variables are often described as probabilistic instead of casual [@Scutari2014].
The information within a node can be either continuous or discrete, and a complete network can contain both continuous and discrete nodes; however, discrete networks are the most commonly used form of BN [@Chen2012]. In discrete networks, conditional probabilities for each child node are allocated for each combination of the possible states in their parent nodes and can be used to assess the strength of a dependency in the network.

@@

Of the 351 participants that were initially recruited for the study, 94 only completed the first time point, and therefore had to be removed from the study as no injury information was obtained for these participants following the first time point. 
To prepare the data for the BN, missing values in the dataset were first imputed. 
Out of the 650 total measurements across all time points in the current study, there were `r n_perc(missingtot$stiffness == 0)` missing muscle stiffness measurements and `r n_perc(missingtot$hrv == 0)` missing heart rate recordings. The missing data were due to technical faults in the data collection equipment and were considered to be missing completely at random. 
A missing rate of 15-20% has been reported to be common in psychological studies, and several techniques are available to handle missing values [@Enders2003; @Lang2014].
In the current study, the _caret_ R package [@Kuhn2008] was used to impute the missing values. 
A bagged tree model using all of the non-missing data was first generated and then used to predict each missing value in the dataset. 
The bagged tree method is a reliable and accurate method for imputing missing values in data and is superior to other commonly used methods such a median imputation [@Kuhn2008].

@@

To obtain the network structure, several steps were taken to ensure that both a theoretically realistic network, and a network that was an appropriate fit to the collected data, was used for inference. 
Prior knowledge about the network structure was included by providing a list of arcs that are always _restricted_ from being in the network (blacklist), and a list of arcs that are always _included_ in the network (whitelist). 
Additionally, there are several scoring functions such as Bayesian Information Criteria (BIC) and Bayesian Dirichlet equivalent uniform (BDeu) that can be used to compare network structures with certain nodes and arcs included or excluded [@Scutari2014]. 
To account for the repeated measures design employed in this study and to maximise the use of the data, pairs of complete cases (e.g., participants who completed T1 + T2, and T2 + T3) were used in a two-time Bayesian network (2TBN) structure [@Murphy2002].
In the 2TBN, variables measured at T2 could depend on variables measured at T1 (e.g., T1 $\rightarrow$ T2) and variables measured at T3 could depend on variables measured at T2 (e.g., T2 $\rightarrow$ T3).
However, arcs were blacklisted between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2 to preserve the order in which data was collected. 
Variables were separated into two groups; "explanatory", for variables that were fixed (e.g., gender), or "independent", for variables that were measured at each time point and could vary during the study.
Independent variable names were suffixed with _1 for time point T, and _2 for time point T+1 (e.g., T1_1 $\rightarrow$ T2_2 and T2_1 $\rightarrow$ T3_2).
Formatting the data in this way meant participants who completed T1 and T2, but did not complete T3, could still be included in the analysis. 
In addition to the blacklisted arcs between T2 $\rightarrow$ T1 and T3 $\rightarrow$ T2, the direction of arcs was restricted between independent variables and explanatory variables (e.g., independent $\rightarrow$ explanatory); however, arcs were not restricted between explanatory $\rightarrow$ independent variables.
Finally, arc direction was restricted between specific nodes within the explanatory variables. 
Arcs from clevel $\rightarrow$ gender, nlebase $\rightarrow$ gender and nlebase $\rightarrow$ ind_team were included in the blacklist, as arcs in these directions did not make logical sense. 
All subsequent models used the same blacklist.

@@

Initially, both negative life events and total life events were included in the network structure, however, the network score was improved when only nlelg or tlelg was included (nleleg = `r round(le[1,2], 2)`, tlelg only = `r round(le[2,2],2)`, tlelg and nlelg = `r round(le[3,2],2)`). 
Additionally, despite strong evidence in the literature that both negative (NLE) and total life event (TLE) stress are related to injury occurrence [@Williams2007], network structures learned using the Tabu search algorithm failed to identify a relationship between NLE and injury or TLE and injury in the data. 
Given that nlelg provided the highest network score, and there is a stronger relationship between negative life events and injury in the literature, an arc was whitelisted between nlelg_1 and injured_1 and nlelg_2 and injured_2 in the final network structure. TLE score was not included in the final structure.

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]`), GDP (`r rstmodel[4,2]`) and I (`r rstmodel[5,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure. 
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

@@

Preliminary network structures also revealed strong dependencies between the same variables at sequential time points. For example, the probability that stiffness_1 and stiffness_2 were both “High”, or both “Low” was approximately 80%. Including the arcs between the same variables from X_1 → X_2 did not provide any theoretically meaningful information to the network structure as the majority of participants would either be in a “Low” or “High” state for each pair of variables in the network. Therefore, these arcs were blacklisted from the network.
To obtain the final network, the appropriate blacklist and whitelists were provided and a Tabu search algorithm identified the remaining structure of the network. 
The final network structure was obtained by averaging 1000 bootstrapped models [@Efron1993] to reduce the impact of locally optimal, but globally suboptimal network learning, and to obtain a more robust model [@Olmedilla2018]. Arcs that were present in at least 30% of the models were included in the averaged model. 
The strength of each arc was determined by the percentage of models that the arc was included in, independent of the arc’s direction. An arc strength of 1 indicated that the arc was always present in the network, with the value decreasing as arcs were found in fewer networks. In the respective study arcs above 0.5 were considered “significant” with arcs below 0.5 and above 0.3 “non-significant” [@Scutari2013]. 
Arcs below 0.3 were not included in the model. The full table of arc strengths for the first and second network are available in supplementary table 1 and supplementary table 3 respectively.

@@

@@

Conditional probability queries (CPQ) were again used to perform inference the network structures.
The Markov blanket of the "injured" node was of particular interest, and the probability of injury was investigated with combinations of variables in the Markov blanket below the mean change, at the mean change and above the mean change.
Initial visual inspection of the network structure also revealed arcs between BIS $\rightarrow$ FFFS and HRV $\rightarrow$ FFFS. 
To investigate this relationship further, random samples were generated for BIS, HRV and FFFS based on the conditional distribution of the nodes included as evidence in the query.
The samples were then examined using a Bayesian linear regression models with the _brms_ R package [@Burkner2017a] to determine the relationship between these nodes. 
Weakly regularising priors (normal prior with mean of 0 and standard deviation of 5) were used for all parameters in the model.

@@

cbind(inj_summary[c(1:3), c(2:4)], inj_summary[c(4:6), c(3:4)]) %>%
  kable(
    .,
    caption = "The number and percentage (\\%) of types of injuries sustained by male and female participants.",
    col.names = c("", "Lower body", "Upper body", "Lower body", "Upper body"),
    align = "lcccc",
    escape = F
  ) %>%
  add_header_above(c(" " = 1, "Female" = 2, "Male" = 2), bold = T) %>%
  footnote(general = "Other included bone, skin and brain injuries.") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE)

@@

```{r fig6, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables. The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
renderGraph(gR)
```

@@

```{r table5, results='asis'}
query1 %>%
  select(c(1, 2, 3)) %>%
  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`

@@

The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".

@@

Table \ref{tab:table11} shows the three highest and lowest probabilities for injury for all variables in the Markov blanket (full results in supplementary table 7).
The combination of 1 SD above the mean change for nlec and stiffness and "High" hours per week and previous injury resulted in the highest probability that an injury would be sustained during the study (`r round(max(query7$prob), digits = 2)`).
In contrast, below average changes in nlec and stiffness combined with "Low" hours per week and no previous injury resulted in the lowest probability of an injury (`r round(min(query7$prob), digits = 2)`).

@@
